### 2015-06-05 Projects/DelCo_2012/Report/DelCo_2012.tex
Added files from personal directories.

Traditionally, fire fighting suppression operations have been conducted from the interior of the structure as a means to reduce water damage, manage the “thermal balance,” and limit fire damage to structures.  In current fire fighter training manuals, it is stated that these operations must be coordinated with the ventilation operations.  However guidance on that coordination has been limited.  Previous research and examinations of line of duty deaths have shown that ventilation events occurring with fire fighters in the structure prior to suppression have led to tragic results.  One means of reducing the possibilities of this occurrence would be to begin to extinguish the fire prior to entering the structure.  In effect, an offensive, exterior fire attack, in which water is directed into the structure from the exterior to cool the fire gases and reduce the heat release rate of the fire, prior to the fire fighters entering the building.  The major concern with this type of operation is the potential harm that might occur to people trapped in the structure or the amount of water damage to the structure.  Therefore, measurements are needed to document the changes of the thermal environment within the structure due to interior and exterior fire suppression tactics.  This report will focus on water distribution and fire suppression experiments conducted to develop a baseline for the interior fire suppression.

@@

NIST FFTG

Objective: To improve the safety and effectiveness of firefighters through measurement science to advance suppression tactics, examine non-traditional means of fire suppression, and transfer of the results to the fire service.

Aid the fire service across the country in the most effective and safe way to fight fires.

The goal is to improve the safety of not only the responding firefighters but the public as well through the use of modern fire suppression tactics for modern fires.

Approach: Real scale, instrumented fire suppression experiments.

This is an STRS project that has been complemented with OA studies funding from: 
Cal Poly 
FDNY 
ISFSI 
NYU

NIST has conducted a significant amount of research examining how ventilation affects the growth and spread of fire within structures and how the air flow to the fire may be controlled to limit or delay the growth of the fire.  The studies have resulted in guidance to the fire service regarding ventilation tactics.  However ventilation tactics alone will not result in the complete extinguishment of the fire, fire suppression with hose streams are needed.

Fire suppression tactics using hose streams also affect the ventilation in a structure and can impact the movement of smoke and heat through a structure as vents are made to advance the line or if ventilation inducing hand line tactics are in practice.  This research addressing the coordination of suppression tactics and the impact on ventilation is needed to complete recommendations on fire control tactics to appropriate standards, education, and training documents.

Test Objectives:

1)	Examine the water distribution for solid streams, straight streams, and fog streams.

2)	Examine the impact of solid streams, straight streams, and fog streams on the cooling of fire gases inside a structure with various flow path configurations, structure configurations, and structure volumes.

3)  Examine the impact of solid streams, straight streams, and fog streams on the

Constants: 
Building Geometry
Fuel Load
Fire Room Conditions at time of water flow
Water Flow Rate – 120 gpm

Variable: 
Smooth Bore vs Narrow Fog

@@

There are different components of uncertainty in the length, mass, temperature, heat flux, gas concentration, differential pressure, gas velocity and heat release rate reported here. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means  [ ]. Type B analysis of systematic uncertainties involves estimating the upper (+ a) and lower (- a) limits for the quantity in question such that the probability that the value would be in the interval (± a) is essentially 100 %. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95 % confidence interval (2σ).  For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

Each length measurement was taken carefully. Length measurements such as the room dimensions, instrumentation array locations and fire apparatus (for example nozzle, sprinkler, or fan) placement were made with a hand held laser measurement device which is has an accuracy of ± 6.0 mm (0.25) over a range of 0.61 m (2.00 ft) to 15.3 m (50.0 ft) [ ] .  However, conditions affecting the measurement, such as levelness of the device, yields an estimated uncertainty of ± 0.5 % for measurements in the 2.0 m (6.6 ft) to 10.0 m (32.8 ft) range.  Steel measuring tapes with a resolution of  ± 0.5 mm (0.02 in) were used to locate individual sensors within a measurement array and to measure and position the furniture. The steel measuring tapes were manufactured in compliance with NIST Manual 44, which specifies a tolerance of ±1.6 mm (0.06 in) for 9.1 m (30 ft) tapes and ±6.4 mm (0.25 in) for 30.5 m (100 ft) tapes [ ] .  Some issues, such as “soft” edges on the upholstered furniture, result in an estimated total expanded uncertainty of  ± 1.0 %.

The load cell used to weigh the fuels prior to the experiments had a range of 0 kg (0 lbs) to 200 kg (440lbs) with a resolution of a 0.05 kg (0.11 lb) and a calibration uncertainty within 1% [] . The expanded uncertainty is estimated to be less than  ± 5 %. 
The standard uncertainty in temperature of the thermocouple wire itself is  ± 2.2 °C at 277 °C and increases to ± 9.5 °C at 871 °C as determined by the wire manufacturer [ ] .  The variation of the temperature in the environment surrounding the thermocouple is known to be much greater than that of the wire uncertainty [ ,  ] .  Small diameter thermocouples were used to limit the impact of radiative heating and cooling.  The estimated total expanded uncertainty for temperature in these experiments is ± 15 %.
In this study, total heat flux measurements were made with water-cooled Schimidt-Bolter gauges.  The manufacturer reports a ± 3 % calibration expanded uncertainty for these devices [ ] . Results from an international study on total heat flux gauge calibration and response demonstrated that the uncertainty of a Schmidt-Boelter gauge is typically ± 8 % [] .

The gas measurement instruments and sampling system used in this series of experiments have been demonstrated an expanded (k = 2) relative uncertainty of ± 1 % when compared with span gas volume fractions [].   Given the non-uniformities and movement of the fire gas environment and the limited set of sampling points in these experiments an estimated uncertainty of ± 12 % is being applied to the results [ ].

Differential pressure reading uncertainty components were derived from pressure transducer
instrument specifications and previous experience with pressure transducers. The transducers
were factory calibrated and the zero and span of each was checked in the laboratory prior to the
experiments yielding an accuracy of ± 1 % [ ] . The total expanded uncertainty was estimated at 10 %.

Bi-directional probes and single thermocouples were used to measure the velocity.  The bi-directional probes used similar pressure transducers as those used for the differential pressure measurements discussed above.   Bare-bead Type K thermocouple are co-located with the probe. A gas velocity measurement study, examining the doorway flow of pre-flashover compartment fires, yielded expanded uncertainty measurements ranging from ± 0.14 to ± 0.22 for bi-directional probes of similar design[] .  The total expanded uncertainty for gas velocity in these experiments w estimated to be  ± 18 %.   
Water Flowrate
Water flowrate was measured with a pressure and flow meter combination shown in Figure x.x. The meter consists of a section of 6.35 cm (2.5 inch) cast aluminum pipe with a 0 – 4.1 MPa (0 - 600 psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800 lpm (1250 gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate.  The manufacturer reports a ± 5 % calibration expanded uncertainty for the flow sensor and ± 3 %  for the pressure sensor [ ] . The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flowrate. The total expanded uncertainty was estimated at ± 10 %.

@@

@@

\begin{sidewaystable}[!ht]
	\centering
	\caption{Fuel masses}
	\begin{tabular}{llcc}
		\hline\noalign{\smallskip}
		Item                         &  Material Description             &  Dimensions (m)            &  Mass (kg)  \\
		\noalign{\smallskip}\hline\noalign{\smallskip}
		3-seat Legacy Sofa         &  491 Brawley                      &  1.96 m x 0.86 m x 0.89 m  &  56.2 kg    \\
		Cushion (left)           &  Cotton w/ inner springs          &                            &  4.06 kg    \\
		Cushions (center)        &  Cotton w/ inner springs          &                            &  4.56 kg    \\
		Cushions (rear)          &  Cotton w/ inner springs          &                            &  4.16 kg    \\
		2-seat New Sofa            &  491 Brawley                      &  1.59 m x 0.91 m x 0.86 m  &  33.3 kg    \\
		Cushion (seat)           &                                   &  0.56 m x 0.61 m x 0.14 m  &  1.6 kg     \\
		Cushion (rear)           &  0.71 m at top, 0.58 m at bottom  &  Varies x 0.53 m x 0.20 m  &  2.0 kg     \\
		3-seat Brown Sofa            &  302 College (front room)         &  2.24 m x 1.02 m x 0.91 m  &  61.0 kg    \\
		2-seat Purple Stripe Sofa    &  302 College (front room)         &  1.57 m x 0.97 m x 0.86 m  &  82.0 kg    \\
		3-seat Black Pine Leaf Sofa  &  302 College (front room)         &  2.08 m x 0.86 m x 0.86 m  &  51.0 kg    \\
		Blue/Brown Chair             &  302 College (front room)         &  0.97 m x 0.81 m x 0.97 m  &  46.4 kg    \\
		2-seat Floral Sofa           &  302 College (rear room)          &  1.68 m x 0.91 m x 0.91 m  &  34.5 kg    \\
		2-seat Blue Plaid Sofa       &  302 College (rear room)          &  1.63 m x 0.97 m x 0.91 m  &  39.8 kg    \\
		3-seat Yellow Sofa           &  302 College (rear room)          &  2.13 m x 0.91 m x 0.71 m  &  48.3 kg    \\
		3-seat Floral Sofa           &  302 College (rear room)          &  2.29 m x 0.91 m x 0.91 m  &  48.0 kg    \\
		3-seat New Sofa              &  289 College (front room)         &  2.13 m x 0.91 m x 0.91 m  &  43.0 kg    \\
		Blue Chair                   &  289 College (middle room)        &  0.91 m x 0.91 m x 0.91 m  &  25.3 kg    \\
		3-seat Floral Sofa           &  289 College (sprinkler room)     &  2.24 m x 1.02 m x 1.02 m  &  81.2 kg    \\
		Floral Chair                 &  289 College (sprinkler room)     &  1.02 m x 1.02 m x 1.02 m  &  36.3 kg    \\
		\noalign{\smallskip}\hline
	\end{tabular}
	\label{tab:Fuel_Masses}
\end{sidewaystable}

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB1}
	\caption{Burn Building Test 1}
	\label{fig:Burn_Building_Test_1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB2}
	\caption{Burn Building Test 2}
	\label{fig:Burn_Building_Test_2}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB3}
	\caption{Burn Building Test 3}
	\label{fig:Burn_Building_Test_3}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB4}
	\caption{Burn Building Test 4}
	\label{fig:Burn_Building_Test_4}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB5}
	\caption{Burn Building Test 5}
	\label{fig:Burn_Building_Test_5}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB6}
	\caption{Burn Building Test 6}
	\label{fig:Burn_Building_Test_6}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB7}
	\caption{Burn Building Test 7}
	\label{fig:Burn_Building_Test_7}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB8}
	\caption{Burn Building Test 8}
	\label{fig:Burn_Building_Test_8}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB9}
	\caption{Burn Building Test 9}
	\label{fig:Burn_Building_Test_9}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB10}
	\caption{Burn Building Test 10}
	\label{fig:Burn_Building_Test_10}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB11}
	\caption{Burn Building Test 11}
	\label{fig:Burn_Building_Test_11}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB12}
	\caption{Burn Building Test 12}
	\label{fig:Burn_Building_Test_12}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB13}
	\caption{Burn Building Test 13}
	\label{fig:Burn_Building_Test_13}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB14}
	\caption{Burn Building Test 14}
	\label{fig:Burn_Building_Test_14}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB15}
	\caption{Burn Building Test 15}
	\label{fig:Burn_Building_Test_15}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB16}
	\caption{Burn Building Test 16}
	\label{fig:Burn_Building_Test_16}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB17}
	\caption{Burn Building Test 17}
	\label{fig:Burn_Building_Test_17}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB18}
	\caption{Burn Building Test 18}
	\label{fig:Burn_Building_Test_18}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB19}
	\caption{Burn Building Test 19}
	\label{fig:Burn_Building_Test_19}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB20}
	\caption{Burn Building Test 20}
	\label{fig:Burn_Building_Test_20}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/BB21}
	\caption{Burn Building Test 21}
	\label{fig:Burn_Building_Test_21}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/ES1}
	\caption{Experimental Structures Test 1}
	\label{fig:Experimental_Structures_Test_1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/ES2}
	\caption{Experimental Structures Test 2}
	\label{fig:Experimental_Structures_Test_2}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/ES3}
	\caption{Experimental Structures Test 3}
	\label{fig:Experimental_Structures_Test_3}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/ES4}
	\caption{Experimental Structures Test 4}
	\label{fig:Experimental_Structures_Test_4}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/ES5}
	\caption{Experimental Structures Test 5}
	\label{fig:Experimental_Structures_Test_5}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/ES6}
	\caption{Experimental Structures Test 6}
	\label{fig:Experimental_Structures_Test_6}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/ES7}
	\caption{Experimental Structures Test 7}
	\label{fig:Experimental_Structures_Test_7}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/ES72}
	\caption{Experimental Structures Test 7.2}
	\label{fig:Experimental_Structures_Test_7.2}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/ES8}
	\caption{Experimental Structures Test 8}
	\label{fig:Experimental_Structures_Test_8}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/ES9}
	\caption{Experimental Structures Test 9}
	\label{fig:Experimental_Structures_Test_9}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/ES10}
	\caption{Experimental Structures Test 10}
	\label{fig:Experimental_Structures_Test_10}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Bars/ES11}
	\caption{Experimental Structures Test 11}
	\label{fig:Experimental_Structures_Test_11}
\end{figure}

@@

The National Institute of Standards and Technology, with the support of the Fire Protection Research
Foundation, the U.S. Department of Homeland Security, and the U.S. Fire Administration conducted a
series of fire experiments to examine the impact of wind on fire spread through a multi-room structure
and examine the capabilities of wind-control devices (WCD) and externally applied water to mitigate the
hazard. The measurements used to examine the impact of the WCDs and the external water application
tactics were heat release rate, temperature, heat flux, and gas velocity inside the structure. Oxygen,
carbon dioxide, carbon monoxide, total hydrocarbons and differential pressures were also measured.
Each of the experiments was recorded with video and thermal imaging cameras. Some of these
measurements are not practical or affordable to make in an acquired structure, hence the need to build a
structure and conduct the experiments within the confines of the NIST Large Fire Facility. These 354
experiments also provided visual documentation of fire phenomena that are not typically observable on
the fire ground.
A limited series of heat release rate experiments were conducted to characterize the fuel load packages
used in wind driven structure experiments. Both the bedroom and the living room contained a fuel load

>>>>

Traditionally, fire fighting suppression operations have been conducted from the interior of the structure as a means to reduce water damage, manage the “thermal balance,” and limit fire damage to structures.  In current fire fighter training manuals, it is stated that these operations must be coordinated with the ventilation operations.  However guidance on that coordination has been limited.  Previous research and examinations of line of duty deaths have shown that ventilation events occurring with fire fighters in the structure prior to suppression have led to tragic results.  One means of reducing the possibilities of this occurrence would be to begin to extinguish the fire prior to entering the structure.  In effect, an offensive, exterior fire attack, in which water is directed into the structure from the exterior to cool the fire gases and reduce the heat release rate of the fire, prior to the fire fighters entering the building.  The major concern with this type of operation is the potential harm that might occur to people trapped in the structure or the amount of water damage to the structure.  Additionally, the type of suppression agent can affect the fire behavior within the structure.  The suppression agent is the main focus of this study to determine whether or not compressed air foam systems provide any significant benefit to the traditional water application.  Therefore, measurements are needed to document the changes of the thermal environment within the structure due to interior and exterior fire suppression tactics over all types of suppression agents.  This report will focus on water distribution and fire suppression experiments conducted to develop a baseline for the interior fire suppression using both water and compressed air foam application.

@@

@@

There are different components of uncertainty in the length, mass, temperature, heat flux, gas concentration, differential pressure, gas velocity and heat release rate reported here. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means  [ ]. Type B analysis of systematic uncertainties involves estimating the upper (+ a) and lower (- a) limits for the quantity in question such that the probability that the value would be in the interval (± a) is essentially 100~\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95~\% confidence interval (2σ).  For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

Each length measurement was taken carefully. Length measurements such as the room dimensions, instrumentation array locations and fire apparatus (for example nozzle, sprinkler, or fan) placement were made with a hand held laser measurement device which is has an accuracy of ± 6.0 mm (0.25) over a range of 0.61 m (2.00 ft) to 15.3 m (50.0 ft) [ ] .  However, conditions affecting the measurement, such as levelness of the device, yields an estimated uncertainty of ± 0.5~\% for measurements in the 2.0 m (6.6 ft) to 10.0 m (32.8 ft) range.  Steel measuring tapes with a resolution of  ± 0.5 mm (0.02 in) were used to locate individual sensors within a measurement array and to measure and position the furniture. The steel measuring tapes were manufactured in compliance with NIST Manual 44, which specifies a tolerance of ±1.6 mm (0.06 in) for 9.1 m (30 ft) tapes and ±6.4 mm (0.25 in) for 30.5 m (100 ft) tapes [ ] .  Some issues, such as “soft” edges on the upholstered furniture, result in an estimated total expanded uncertainty of  ± 1.0~\%.

The load cell used to weigh the fuels prior to the experiments had a range of 0 kg (0 lbs) to 200 kg (440lbs) with a resolution of a 0.05 kg (0.11 lb) and a calibration uncertainty within 1~\% [] . The expanded uncertainty is estimated to be less than  ± 5 %. 
The standard uncertainty in temperature of the thermocouple wire itself is  ± 2.2 °C at 277 °C and increases to ± 9.5 °C at 871 °C as determined by the wire manufacturer [ ] .  The variation of the temperature in the environment surrounding the thermocouple is known to be much greater than that of the wire uncertainty [ ,  ] .  Small diameter thermocouples were used to limit the impact of radiative heating and cooling.  The estimated total expanded uncertainty for temperature in these experiments is ± 15 %.
In this study, total heat flux measurements were made with water-cooled Schimidt-Bolter gauges.  The manufacturer reports a ± 3~\% calibration expanded uncertainty for these devices [ ] . Results from an international study on total heat flux gauge calibration and response demonstrated that the uncertainty of a Schmidt-Boelter gauge is typically ± 8~\% [] .

The gas measurement instruments and sampling system used in this series of experiments have been demonstrated an expanded (k = 2) relative uncertainty of ± 1~\% when compared with span gas volume fractions [].   Given the non-uniformities and movement of the fire gas environment and the limited set of sampling points in these experiments an estimated uncertainty of ± 12~\% is being applied to the results [ ].

Differential pressure reading uncertainty components were derived from pressure transducer
instrument specifications and previous experience with pressure transducers. The transducers
were factory calibrated and the zero and span of each was checked in the laboratory prior to the
experiments yielding an accuracy of ± 1~\% [ ] . The total expanded uncertainty was estimated at 10~\%.

Bi-directional probes and single thermocouples were used to measure the velocity.  The bi-directional probes used similar pressure transducers as those used for the differential pressure measurements discussed above.   Bare-bead Type K thermocouple are co-located with the probe. A gas velocity measurement study, examining the doorway flow of pre-flashover compartment fires, yielded expanded uncertainty measurements ranging from ± 0.14 to ± 0.22 for bi-directional probes of similar design[] .  The total expanded uncertainty for gas velocity in these experiments w estimated to be  ± 18 %.   
Water Flowrate
Water flowrate was measured with a pressure and flow meter combination shown in Figure x.x. The meter consists of a section of 6.35 cm (2.5 inch) cast aluminum pipe with a 0 – 4.1 MPa (0 - 600 psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800 lpm (1250 gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate.  The manufacturer reports a ± 5~\% calibration expanded uncertainty for the flow sensor and ± 3~\%  for the pressure sensor [ ] . The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flowrate. The total expanded uncertainty was estimated at ± 10 %.

@@

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Pictures/PalletFuelLoad}
	\caption{Pallet Fuel Load}
	\label{fig:Pallet_Fuel_Load}
\end{figure}

@@

\begin{sidewaystable}[!ht]
	\centering
	\caption{Fire Suppression Fuel masses}
	\renewcommand{\tabcolsep}{1pt}
	\begin{tabular}{lllllll}
		\hline\noalign{\smallskip}
		Item               & Component		& Quantity		&  Material Description             							&  Dimensions (m)            	&  Mass (kg)  		& Total Mass (kg) \\
		\noalign{\smallskip}\hline\noalign{\smallskip}
		2-seat New Sofa    &				& 2				& 85~\% Urethane Foam, 15~\% Polyester Fiber and Wood Frame 	&  1.68 L X 0.90 D X 0.86 H  	&  72.4    			& 144.8 \\
		           	 	   & \multirow{2}{*}{{Mattress}}	    & \multirow{2}{*}{{1 per sofa}}	& \multirow{2}{*}{\parbox{5cm} {Cover: 52~\% Polypropylene, 48~\% Polyester, Filling: 66~\% Polyester Fiber Pad, 34~\% Polyester Fiber  Batting}} & \multirow{2}{*}{{1.83 L X 1.21 W X 0.14 H}}  & \multirow{2}{*}{{14.6}}     \\
		           	 	   &	            &               &                                                          		&                            	&             		& \\[.25cm]
			           	   & Seat Cushion	& 2 per sofa	& 85~\% Urethane Foam, 15~\% Polyester Fiber   				    &  0.69 L X 0.58 D X 0.14 H  	&  2.2     			& \\
		Blue Chair         &				& 2				& 85~\% Urethane Foam, 15~\% Polyester Fiber and Wood Frame   	&  0.81 L X 0.75 D X 0.91 H  	&  17.4    			& 34.8 \\
		    			   & Seat Cushion   & 1 per chair   & 85~\% Urethane Foam, 15~\% Polyester Fiber					&  0.51 L X 0.57 D X 0.14 H  	&  1.7    			& \\
		Blue/Green Chair   & 				& 2				& 85~\% Urethane Foam, 15~\% Polyester Fiber and Wood Frame     &  0.77 L X 0.71 D X 0.89 H  	&  16    			& 32 \\
		                   & Seat Cushion	& 1 per chair	& 85~\% Urethane Foam, 15~\% Polyester Fiber		            &  0.55 L X 0.53 D X 0.15 H  	&  1.6    			& \\
		Paneling           &				& 8 panels		& 3.2 mm thick medium density fiberboard                   		&  2.44 L X 1.22 W X 3.2mm thick&  9    			& 72 \\
		Carpeting          &				& 22.26 m$^2$	& 100~\% PET Polyester                  						&  3.68 L X 6.05 W  			&  1.44 kg/m$^2$	& 32.2 \\
		Carpet Padding     &				& 22.26 m$^2$	& Polyurethane                 	 								&  3.68 L X 6.05 W X 11mm thick &  0.89 kg/m$^2$    & 19.8 \\
		                   &                &               &                                                               &                               & Total Mass        & 335.6 \\
		\noalign{\smallskip}\hline
	\end{tabular} 
	\label{tab:Fire_Suppression_Fuel_Masses}
\end{sidewaystable}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Pictures/SleeperSofa}
	\caption{Sleeper Sofa}
	\label{fig:Sleeper_Sofa}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Pictures/Mattress}
	\caption{Mattress}
	\label{fig:Mattress}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=6in]{../Figures/Pictures/BlueChair}
	\caption{Blue Chair}
	\label{fig:Blue_Chair}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB1}
	\caption{Burn Building Test 1}
	\label{fig:Burn_Building_Test_1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB2}
	\caption{Burn Building Test 2}
	\label{fig:Burn_Building_Test_2}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB3}
	\caption{Burn Building Test 3}
	\label{fig:Burn_Building_Test_3}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB4}
	\caption{Burn Building Test 4}
	\label{fig:Burn_Building_Test_4}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB5}
	\caption{Burn Building Test 5}
	\label{fig:Burn_Building_Test_5}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB6}
	\caption{Burn Building Test 6}
	\label{fig:Burn_Building_Test_6}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB7}
	\caption{Burn Building Test 7}
	\label{fig:Burn_Building_Test_7}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB8}
	\caption{Burn Building Test 8}
	\label{fig:Burn_Building_Test_8}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB9}
	\caption{Burn Building Test 9}
	\label{fig:Burn_Building_Test_9}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB10}
	\caption{Burn Building Test 10}
	\label{fig:Burn_Building_Test_10}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB11}
	\caption{Burn Building Test 11}
	\label{fig:Burn_Building_Test_11}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB12}
	\caption{Burn Building Test 12}
	\label{fig:Burn_Building_Test_12}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB13}
	\caption{Burn Building Test 13}
	\label{fig:Burn_Building_Test_13}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB14}
	\caption{Burn Building Test 14}
	\label{fig:Burn_Building_Test_14}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB15}
	\caption{Burn Building Test 15}
	\label{fig:Burn_Building_Test_15}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB16}
	\caption{Burn Building Test 16}
	\label{fig:Burn_Building_Test_16}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB17}
	\caption{Burn Building Test 17}
	\label{fig:Burn_Building_Test_17}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB18}
	\caption{Burn Building Test 18}
	\label{fig:Burn_Building_Test_18}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB19}
	\caption{Burn Building Test 19}
	\label{fig:Burn_Building_Test_19}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB20}
	\caption{Burn Building Test 20}
	\label{fig:Burn_Building_Test_20}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/BB21}
	\caption{Burn Building Test 21}
	\label{fig:Burn_Building_Test_21}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/ES1}
	\caption{Experimental Structures Test 1}
	\label{fig:Experimental_Structures_Test_1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/ES2}
	\caption{Experimental Structures Test 2}
	\label{fig:Experimental_Structures_Test_2}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/ES3}
	\caption{Experimental Structures Test 3}
	\label{fig:Experimental_Structures_Test_3}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/ES4}
	\caption{Experimental Structures Test 4}
	\label{fig:Experimental_Structures_Test_4}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/ES5}
	\caption{Experimental Structures Test 5}
	\label{fig:Experimental_Structures_Test_5}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/ES6}
	\caption{Experimental Structures Test 6}
	\label{fig:Experimental_Structures_Test_6}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/ES7}
	\caption{Experimental Structures Test 7}
	\label{fig:Experimental_Structures_Test_7}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/ES72}
	\caption{Experimental Structures Test 7.2}
	\label{fig:Experimental_Structures_Test_7.2}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/ES8}
	\caption{Experimental Structures Test 8}
	\label{fig:Experimental_Structures_Test_8}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/ES9}
	\caption{Experimental Structures Test 9}
	\label{fig:Experimental_Structures_Test_9}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/ES10}
	\caption{Experimental Structures Test 10}
	\label{fig:Experimental_Structures_Test_10}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=4in]{../Figures/Bars/ES11}
	\caption{Experimental Structures Test 11}
	\label{fig:Experimental_Structures_Test_11}
\end{figure}

@@

Smoke is Fuel. A ventilation limited (fuel rich) condition had developed prior to the failure of
the window. Oxygen depleted combustion products, containing carbon dioxide, carbon monoxide and
unburned hydrocarbons filled the rooms of the structure. Once the window failed, the fresh air provided
the oxygen needed to sustain the transition through flashover, which caused a significant increase in heat
release rate.
Venting does not always equal cooling. In this experiment, post ventilation temperatures and
heat fluxes all increased, due to the ventilation induced flashover.
Fire induced flows. Velocities within the structure exceeded 5 m/s (11 mph), just due to the fire
growth and the flow path that was set-up between the window opening and the corridor vent.
Avoid the flow path. The directional nature of the fire gas flow was demonstrated with thermal
conditions, both temperature and heat flux, which were twice as high in the “flow” portion of the
corridor as opposed to the “static” portion of the corridor. Thermal conditions in the
flow path were not consistent with firefighter survival.

The fuel load in the structure was the same for all of the experiments. Each of these experiments
demonstrated a rapid transition to untenable conditions in the corridor, even for a firefighter in full PPE.
